Vagueness, Noise, and Communication

### 2014-10-07 paper/01_intro.tex
Small corrections to background section.

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi-languages. Intuitively speaking, a
Voronoi language is a pair of sender and receiver strategies, such
that the sender strategy (quasi-)partitions the state space into
convex categories, while the receiver's interpretations are the
central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
an example for such a Voronoi-language for a sim-max game with
$\States = [0,1]$, two messages, and a flat prior distribution that
selects all states equally likely. The sender's strategy (a function
from states to messages; more below) uses one message exclusively for
all points in the lower interval and another for all points in the
upper interval. The receiver's interpretations of messages are the
central points in the respective intervals.

@@

This result is particularly interesting, because it demonstrates that
signaling can impose orderly categories on a metric space, without
that being the ulterior purpose of it all. Sender and receiver can be
distinct entities, whose purpose it is to communicate effectively
about the actual state. In that case, evolving Voronoi-languages would
explain why linguistic categories are well-behaved and orderly in the
way they appear to be. More abstractly, sender and receiver can also
be thought of as distinct modules in a single system, where the first
module that must discretize the information that it is fed into a
small sample of, suggestively, category labels to pass to a second
system that tries to decode the original information. In this case,
evolving Voronoi-languages would explain why conceptual categories are
well-behaved and orderly in the way that they appear to be
\citep[see][for more on this latter
interpretation]{OConnor2013:Evolving-Percep}.

Seen in this light, sim-max games appear to have the potential of
providing a foundation to approaches in cognitive semantics that rely
on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has
prominently argued that natural categories are convex regions in a
suitable conceptual space. If the conceptual space has a suitable
metric, then it is possible to think of these convex categories as
derived from a set of prototypes. Fixing a set of prototypes, we
consider the category corresponding to each prototype $p$ as the set
of points that are more similar to $p$ than to any other
\citep[e.g.][]{OkabeBoots2000:Spatial-Tessell}. In this way,
\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorization system can be obtained: storing the prototypes lets us
recover the categories without having to store each categories
extension. However, what is left unexplained so far, is where the
prototypes come from, and why we would not see just any distribution
of prototypes as an equally efficient classification system. This is
where sim-max games, might chip in a principled approach to deriving
in an independent way, not only convex categories, but also
prototypical exemplars belonging to them.

@@

The brief outline of an approach to conceptual categorization using
sim-max games leaves many problems unaddressed. One of them is that
natural categories for continuously variable stimuli, like shades of
color, pitch heights, spatial dimension and the like, do not have
unique prototypes and clear category boundaries. We would like to
account for the possibility of vagueness, in particular the following
criteria
\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re,Smith2008:Vagueness-and-D}:
(i) clear positive examples of a vague category should show a
gradient, perhaps smooth, transition to negative examples to
accommodate also ``higher-order vagueness'' ; (ii) prototypes should
likewise be gradient regions, peaking at the center of the vague
category they represent.

@@

An alternative approach is taken by
\citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o} who show how the above desiderata
can be met by evolving strategies in sim-max games for certain types
of solution concepts, that can be divided into micro- and macro-level
approaches. Micro-level approaches look at adaptive behavior of
individual agents. Usually, changes in the behavioral dispositions of
agents occur after every single interaction. In contrast, macro-level
approaches outline more abstract, aggregate dynamics, happening in a
huge population of agents, or otherwise abstracting from seemingly
irrelevant detail. Usually, a macro-level dynamic captures changes of
frequencies of behavioral types in the population over time.

>>>>

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi-languages. Intuitively speaking, a
Voronoi language is a pair of sender and receiver strategies, such
that the sender strategy (quasi-)partitions the state space into
convex categories, while the receiver's interpretations are the
central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
an example of such a Voronoi-language for a sim-max game with
$\States = [0,1]$, two messages, and a flat prior distribution that
selects all states equally likely. The sender's strategy (a function
from states to messages; more below) uses one message exclusively for
all points in the lower half of the interval and another for all points in the
upper half. The receiver's interpretations of messages are the
central points in the respective intervals.

@@

This result is particularly interesting, because it demonstrates that
signaling can impose orderly categories on a metric space, without
that being the ulterior purpose of it all. Sender and receiver can be
distinct entities, whose purpose is to communicate effectively
about the actual state. In that case, evolving Voronoi-languages would
explain why linguistic categories are well-behaved and orderly in the
way they appear to be. More abstractly, sender and receiver can also
be thought of as distinct modules in a single system, where the first
module that must discretize the information that it is fed into a
small sample of, suggestively, category labels to pass to a second
system that tries to decode the original information. In this case,
evolving Voronoi-languages would explain why conceptual categories are
well-behaved and orderly in the way that they appear to be
\citep[see][for more on this latter
interpretation]{OConnor2013:Evolving-Percep}.

Seen in this light, sim-max games appear to have the potential of
providing a foundation to approaches in cognitive semantics that rely
on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has
prominently argued that natural categories are convex regions in a
suitable conceptual space. If the conceptual space has a suitable
metric, then it is possible to think of these convex categories as
derived from a set of prototypes. Fixing a set of prototypes, we
consider the category corresponding to each prototype $p$ as the set
of points that are more similar to $p$ than to any other
\citep[e.g.][]{OkabeBoots2000:Spatial-Tessell}. In this way,
\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorization system can be obtained: storing the prototypes lets us
recover the categories without having to store each categories
extension. However, what is left unexplained so far, is where the
prototypes come from, and why we would not see just any distribution
of prototypes as an equally efficient classification system. This is
where sim-max games might chip in a principled approach to deriving,
in an independent way, not only convex categories but also
prototypical exemplars belonging to them.

@@

The brief outline of an approach to conceptual categorization using
sim-max games leaves many problems unaddressed. One of them is that
natural categories for continuously variable stimuli, like shades of
color, pitch heights, spatial dimension and the like, do not have
unique prototypes and clear category boundaries. We would like to
account for the possibility of vagueness, in particular the following
criteria
\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re,Smith2008:Vagueness-and-D}:
(i) clear positive examples of a vague category should show a
gradient, perhaps smooth, transition to negative examples to
accommodate also higher-order vagueness; (ii) prototypes should
likewise be gradient regions, peaking at the center of the vague
category they represent.

@@

An alternative approach is taken by
\citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o} who show how the above desiderata
can be met by evolving strategies in sim-max games for certain types
of solution concepts which can be divided into micro- and macro-level
approaches. Micro-level approaches look at adaptive behavior of
individual agents. Usually, changes in the behavioral dispositions of
agents occur after every single interaction. In contrast, macro-level
approaches outline more abstract, aggregate dynamics, happening in a
huge population of agents, or otherwise abstracting from seemingly
irrelevant detail. Usually, a macro-level dynamic captures changes of
frequencies of behavioral types in the population over time.
